Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-01-09 main_backup.tex
update figures, fix code issues

In many fields it is common practice to forecast transformed quantities (see e.g. \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology). While the goal of the transformations is usually to improve accuracy of the predictions, they can also be used to enhance and complement the evaluation process. 
In this paper we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation. Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score. Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).

@@

Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets. The change in rankings of aggregate scores is not mainly driven by the kinds of changes in model rankings for single forecasts just discussed. Rather, average scores will be most influenced by the order of magnitude of scores for different forecast targets across time, location and target type. On the natural scale, large observations will account for the majority of the overall score but will not dominate average CRPS values obtained after log transformation in the same way. Depending on the relationship between mean and variance of the forecast target, a log-transformation, however, may also lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}. A different transformation like the square-root may then be appropriate to stabilise the variance effectively and make scores independent of the mean of the target quantity.

@@

%Whether forecasts of small or large targets receive higher scores than those for targets with larger quantities, depends on the relationship between the mean and the variability of the quantity of interest (see Section \ref{sec:methods:vst}. For the simulated count data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$ (which for large $\mu$ is dominated by the quadratic term), and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).

@@

In practice, one issue with transformations, such as the log transform, is that they may not be readily applicable to negative numbers or zero values, which need to be removed or handled otherwise. Small discrete values can pose an additional problem as it may be difficult to reliably evaluate relative errors. 
One common approach is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm. This represents a strictly monotonic transformation and therefore still preserves the propriety of the resulting score. The choice of the number to add influences scores and rankings, as measures of relative errors shrink when adding a constant to the forecast and the observation. Adding a number to the forecasts and observations changes scores for forecasts of small quantities most, as relative errors shrink more the smaller the original value before adding a constant. In principle, it would also be possible to exploit the difference in how scores are affected in order to equalise scores for forecasts of quantities across different scales by adding a larger number (e.g. 100 or 10,000) if average scores are dominated too strongly by forecasts for small quantities.

@@

The log-transformation is only one of many transformation that may be useful and appropriate in an epidemiological context. One easy transformation is to divide all forecasts by the population in a location, in order to obtain forecasts of standardised incidences per e.g. 100,000 inhabitants. A second transformation was mentioned above, namely converting forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value. One could also think about applying a Box-Cox \citep{boxAnalysisTransformations1964} or a square-root transformation in order to stabilise the variance of the forecasts. Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week multiplicative growth rates, or alternatively, to ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly. This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluate the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}. This, unfortunately is not feasible in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available. What makes this approach of transforming forecasts before scoring particularly interesting is the possibility to construct composite scores as a weighted sum of scores based on different transformations. This would allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.

Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policy makers by providing scoring rules that better reflect what forecast consumers care about. Potentially, the variance stabilising time-series forecasting literature may be a useful source of transformations for various forecast settings. We have showed that the log-transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions. While it is commonly accepted that multiple proper scoring rules should usually be considered when comparing forecasts, we think this should be supplemented by considering different transformations of the data to obtain a richer picture of model performance. More work needs to be done to better understand the effects of applying transformations more comprehensively and in different contexts, and how it might affect decision-making.

>>>>

In many fields it is common practice to forecast transformed quantities (see e.g. \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology). While the goal of the transformations is usually to improve accuracy of the predictions, they can also be used to enhance and complement the evaluation process. 
In this paper we argue that the aforementioned issues with evaluating forecasts based on measures of absolute error can be addressed by transforming the forecasts and observations prior to scoring. Transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score. Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).

@@

Overall model rankings would be expected to differ even more when scores are averaged across multiple forecasts or targets. The change in rankings of aggregate scores is not mainly driven by the kinds of changes in model rankings for single forecasts just discussed. Rather, average scores will be most influenced by the order of magnitude of scores for different forecast targets across time, location and target type. On the natural scale, large observations will account for the majority of the absolute score but will not dominate relative scores on the log scale in the same way. Depending on the relationship between mean and variance of the forecast target, a log-transformation, however, may also lead to systematically larger scores assigned to small forecast targets, as illustrated in Figure \ref{fig:SIM-wis-state-size-mean}. A different transformation like the square-root may then be appropriate to stabilise the variance effectively and make scores independent of the mean of the target quantity.

@@

Whether forecasts of small or large targets receive higher scores than those for targets with larger quantities, depends on the relationship between the mean and the variability of the quantity of interest (see Section \ref{sec:methods:vst}. For the simulated count data, log predictions for small quantities received on average higher scores than forecasts for large quantities if the mean and the variance grew at the same rate $\sigma^2 = \mu$, about equal scores when the variance grew at a rate of $\sigma^2 = \mu + \mu^2$ (which for large $\mu$ is dominated by the quadratic term), and smaller scores when the variance grew faster than that (Figure \ref{fig:SIM-wis-state-size-mean}).

@@

In practice, one issue with transformations, such as the log transform, is that they may not be readily applicable to negative numbers or zero values. When using a log-transformation, it may also be difficult to reliably evaluate relative errors when dealing with small discrete values. In an epidemiological setting that involves count data, negative values should likely be omitted entirely. One potential solution to deal with zeroes and small values is to also exclude these from the ongoing analysis. An alternative is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm as is common practice when log-transforming counts. This represents a strictly monotonic transformation and therefore still preserves the propriety of the resulting score. The choice of the number to add influences scores and rankings, as measures of relative errors shrink when adding a constant to the forecast and the observation. Adding a number to the forecasts and observations changes scores for forecasts of small quantities most, as relative errors shrink more the smaller the original value before adding a constant. In principle, it would also be possible to exploit the difference in how scores are affected in order to equalise scores for forecasts of quantities across different scales by adding a larger number (e.g. 100 or 10,000) if average scores are dominated too strongly by forecasts for small quantities.

@@

The log-transformation is only one of many transformation that may be useful and appropriate in an epidemiological context. One easy transformation is to divide all forecasts by the population in a location, in order to obtain forecasts of standardised incidences per e.g. 100,000 inhabitants. A second transformation was mentioned above, namely converting forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last observed value. One could also think about applying a Box-Cox or a square-root transformation in order to stabilise the variance of the forecasts. Another promising transformation would be to divide each forecast by the forecast of the previous week (and observations by the observation of the previous week), in order to obtain forecasts for week-to-week multiplicative growth rates, or alternatively, to ask forecasters to provide estimates of the weekly relative change applied to the latest data and subsequent forecast points directly. This would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluate the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}. This, unfortunately is not feasible in the context of forecasts stored as predictive quantiles, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate, but may be an interesting approach if predictive samples are available. What makes this approach of transforming forecasts before scoring particularly interesting is the possibility to construct composite scores as a weighted sum of scores based on different transformations. This would allow forecast consumers to assign explicit weights to different qualities of the forecasts they might care about.

Exploring transformations is a promising avenue for future work that could help bridge the gap between modellers and policy makers by providing scoring rules that better reflect what forecast consumers care about. Potentially the variance stabilising time-series forecasting literature may be a useful source of transformations for various forecast settings. We have showed that the log-transformation can lead to significant changes in the relative rankings of models against each other, with potentially important implications for decision-makers who rely on the knowledge of past performance to make a decision about which forecasts should inform future decisions. While it is commonly accepted that multiple proper scoring rules should usually be considered when comparing forecasts, we think this should be supplemented by considering different transformations of the data to obtain a richer picture of model performance. More work needs to be done to better understand the effects of applying transformations more comprehensively and in different contexts, and how it might affect decision-making.
